---
title: "Data Analysis 2 - Assigment 2"
subtitle: "Modelling Fish Weights based on Length, Height, Width and Species" 
date: 2021-01-01
author: "Yuri Almeida Cunha"
output: 
prettydoc::html_pretty:
    theme: architect
    highlight: vignette

---

```{r, echo= FALSE , include = FALSE }

# Load necessary packages

# Data manipulation
library(tidyverse)

# Correlation Matrix
library(Hmisc)

# Different graph thenmes
library(ggthemes)

# For scaling ggplots
library(scales)

# Estimate robust SE
library(lmtest)
library(estimatr)

# Compare models with robust SE
library(texreg)

# Display formatted tables
library(knitr)

# Comparing Models
library(huxtable)
library(jtools)

# print htmltables
library(shiny)


# Clear variable environmental memory
rm(list=ls())

# Call the data from github
my_url <- "https://raw.githubusercontent.com/yurialmeida5/DA2_Final_Project/master/data/clean/fish_data_clean.csv"

# Upload the cleaned data to a dataframe

fish_data <- read_csv( my_url )


```

## Introduction

The objective of the analysis is to develop a multiple regression model predicting the **Weight** in grams **(dependent variable)** of a fish using its **physical dimensions**:
Height, Width, Lengths (1,2,3) and its specie type. **(Explanatory Variables)**.
&nbsp;

The **Lengths** are continuous variables measured in **centimeters (cm)**.    

The **Height** is the maximum height as % of **Length3 (Height =  Height/Length3 * 100)**  

The **Width** is the maximum height as % of **Length3 (Width  =  Width/Length3 * 100)** 

There are 7 species in the sample, with different numbers of observations of each.  
The variable **Species** was transformed into dummy variables named from **(Specie1 to Specie7)**  
The data set is sorted by species from 1 to 7 and then approximately by length within a species.  

## Variables Descriptive Analysis 

### Summary Statistics Table (without Species Dummy Variables)

```{r, echo= FALSE}

table_summary <- fish_data %>% 
  select(-c(Species1,Species2,Species3,Species4, Species5, Species6, Species7)) %>% 
  summary() 

kable(table_summary)

```

### Check the histograms distribution (without Species Dummy Variables)

```{r, echo= FALSE, warning=FALSE, message=FALSE }

fish_data %>% select(-c(Species1,Species2,Species3,Species4, Species5, Species6, Species7)) %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free") +
  geom_histogram()+
  scale_fill_wsj() 

```

Considering the analysis of the graphs and the summary statistics, it's possible to visualize that: 

Whether **Height and Width** are closer to the regular **normal distribution**, the **Length(1,2,3) and Weight** variables are slightly skewed forming **Right-Tails** with the presence of some outliers.  
Taking that into account, it's possible to apply some **log transformations** while the models are build, specially at the dependent variable **Weight** and the explanatory **Length**.  
Variable **Sex** shouldn't be used for analysis because it has a good amount of **missing values** that may impact the model.

## Variables Correlation Matrix

```{r, echo= FALSE}

# Checking correlation between variables 

corr_matrix <- rcorr(as.matrix(select(
  fish_data , -c(Species1,Species2,Species3,Species4, Species5, Species6, Species7, Sex))))

# Correlation Matrix Values
kable(corr_matrix$r)

```

As the correlation **results** between the Length's showed pretty much **closer to 1** with significant p-values:

** Length1 ~ Length3: r = 0.99 and p = 0.000.  
** Length2 ~ Length3: r = 0.99 and p = 0.000.  
** Length1 ~ Length2: r = 1.00 and p = 0.000.  

Having the three of them at the same regression model may lead to errors and won't produce significant better adjustments. With that, the variables **Length1 and Length2 won't be considered at any analysis**. 
Only the variable **Length3**, the total size of the fish, is going to be used as the **Final_Length measurement**.

```{r, echo=FALSE, include=FALSE}

fish_data <- fish_data %>% 
  select(-c(Length1,Length2,Sex))

fish_data <- rename(fish_data, Final_Length = Length3)


```


## Weight and Length

### ln(Weight) ~ ln(Final_Length) 

```{r, echo=FALSE, message=FALSE , out.width='50%'}

ggplot( fish_data , aes(x = Final_Length , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "ln(Length) (cm)", y = "ln(Weight) (grams)") +
  scale_y_continuous( trans = log_trans()) +
  scale_x_continuous(trans = log_trans(), breaks = c(10,20,30,40,50,60, 70))+
  ggtitle("ln(Weight) ~ ln(Length)") +
  theme(plot.title = element_text(hjust = 0.5))


```

According to the results obtained from the graphs and model comparison (please review the following **Appendix section** for more detailed information), the **log-log regression** shows not only a **better fit**, but also an **extraordinary R-squared** adjusted with a really **significant p-value**.
This is supported by the **Right-Tailed Distributions Patterns** encountered both in **Weight** and **Length**.  
The variable **Length itself** could be used **alone** for the **predictions**.

Considering that, the first regression model for analysis is:  
**Reg1:** *ln(Weight) ~ ln(Final_Length)*

## Weight and Height

### ln(Weight) ~ Height

```{r, echo=FALSE, message=FALSE , out.width='50%'}

ggplot( fish_data , aes(x = Height , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Height (%)", y = "ln(Weight) (grams)") +
  scale_y_continuous(trans = log_trans())+
  ggtitle("ln(Weight) ~ Height") +
  theme(plot.title = element_text(hjust = 0.5))


```


The analysis of the graphs and the comparisons between the **models with and without** the log-transformations (please review the following **Appendix section** for more detailed information), didn't show **any significant statistic difference**. The **pattern** expressed in both graphs are quite **similar**, and the results obtained in terms of **R-squared** are pretty **close** as well. 
With that, due to the **ease and more tangible interpretation** of the results, the model **chosen** was the **log-lvl** one.

Another important aspect to highlight is the fact that the pattern observed can possibly fit a **quadratic or cubic polynomial models**. 
Although the cubic and quadratic models showed a slightly better fit than the simple log-lvl, **the interpretation is much easier under the simple model**. 
The differences aren't that significant enough in terms R-Squared as well.

Considering that, the second regression model for analysis is:  
**Reg2:** *ln(Weight) ~ ln(Final_Length) + Height*

## Weight and Width

### ln(Weight) ~ Width

```{r, echo=FALSE, message=FALSE , out.width='50%'}

ggplot( fish_data , aes(x = Height , y = Width)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Width (%)", y = "ln(Weight) (grams)") +
  scale_y_continuous(trans = log_trans())+
  ggtitle("ln(Weight) ~ Width") +
  theme(plot.title = element_text(hjust = 0.5))


```


The analysis of the graphs and the comparisons between the **models with and without** the log-transformations (please review the following **Appendix section** for more detailed information), didn't show **any significant statistic difference**. The **pattern** expressed in both graphs are quite **similar**, and the results obtained in terms of **R-squared** are pretty **close** as well. 
With that, due to the **ease and more tangible interpretation** of the results, the model **chosen** was the **log-lvl** one.

Another important aspect to highlight is the fact that the pattern observed can possibly fit a **quadratic or cubic polynomial models**. 
Although the cubic and quadratic models showed a slightly better fit than the simple log-lvl, **the interpretation is much easier under the simple model**. 
The differences aren't that significant enough in terms R-Squared as well.

Considering that, the second regression model for analysis is:  
**Reg5:** *ln(Weight) ~ ln(Final_Length) + Height + Width*

## Weight and Species

Considering what was done so far, the best possible model so far is represented by reg5. However, it's still missing the impact of the species variables. 
Another regression model can actually be created for that situation:

**Reg8:** *ln(Weight) ~ ln(Final_Length) + Height + Width + as.factor(Species)*

```{r, echo=FALSE, include=FALSE}

fish_data <- fish_data %>% mutate(ln_Weight = log(Weight),
                    ln_Final_Length = log(Final_Length))

reg1 <- lm_robust( ln_Weight  ~ ln_Final_Length  , data = fish_data , se_type = "HC2" )


reg2 <- lm_robust( ln_Weight  ~ ln_Final_Length + Height , data = fish_data , se_type = "HC2")

reg5 <- lm_robust( ln_Weight  ~ ln_Final_Length + Height + Width , data = fish_data , se_type = "HC2" )

reg8 <- lm_robust( ln_Weight  ~ ln_Final_Length + Height + Width + Species1 + Species2 + Species3 + Species4 + Species5 + Species6 , data = fish_data , se_type = "HC2" )

```

```{r, echo=FALSE , out.width='50%'}

as_hux(export_summs(reg1, reg2, reg5, reg8, 
                    model.names = c('reg1', 'reg2', 'reg5', 'reg8')), width(0.4), position(wrapright), caption = 'Weight ~ Length + Width + Height + Species')


```

Adding the dummy variables created for species, the final model, **reg8**, shows itself really representative with **higher fit** and **prediction capabilities**. 
Although the R-squared **didn't present any substantive statistical** difference from the previous models, the **p-values were highly significant**. It means that the model explains well the dataset chosen, using a good number of really representative variables.

For all of that, the **reg8** is going to be the model **used for predictions**. (In case you are reading the pdf, please check the **final table resume** at page number 15 - **Appendix**)

## Predictions

```{r, echo=FALSE, include=FALSE}

set.seed(123)

# Get the predicted y values from the model
fish_data$reg8_y_pred <- reg8$fitted.values

# Calculate the errors of the model
fish_data$reg8_res <- fish_data$ln_Weight - fish_data$reg8_y_pred 


```

```{r, echo=FALSE, out.width='50%' }

# Create: y_hat-y plot
ggplot( data = fish_data ) +
  geom_point (aes( x = reg8_y_pred , y = ln_Weight ) ,  color='red')+
  geom_line( aes( x = ln_Weight , y = ln_Weight ) , color = 'navyblue' , size = 1.5 )+
  labs( x = 'Predicted fish weights', y = 'Actual fish Weights')+
  ggtitle("y_hat - y plot") +
  theme(plot.title = element_text(hjust = 0.5))

```

As expected the model **predicted very well** the values for **Weight**. The **residuals** are pretty low and the graph shows small discrepancy between the actual and the predicted values.
It`s important to mention that that the **residuals were evaluated in terms of the log transformations, as well the values predicted**. The **conversion** from **logarithmic Weight to absolute values should be done carefully**, considering some possible error bias. 
The **variables** presented themselves significant **(lower p-values)** and important for the model description. The **R-squared value** is really **representative** and generates a great understanding of how the dependent variable **Weight is changing in the dataset**.

## Conclusion

The **main goal** of this project, create a reasonable model that could predict the fish Weight based on its physical characteristics, was achieved **successfully**. Considering the **high R-Squared** and **lower p-values**, the model was build with a **strong comprehend of the dataset** structure and patterns generating on that way really **accurate predictions**.

Although it's **impossible** to infer any **causality**, this model seems a **good income** for **further studies** in the area.

\newpage

## Appendix

## Correlation Matrix

```{r, echo=FALSE}


# Correlation Matrix Values
corr_matrix$r

corr_matrix$P

```



### Weight and Length

```{r, echo=FALSE}

# 1.1) Weight ~ Final_Length: level-level model without scaling

ggplot( fish_data , aes(x = Final_Length , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Length (cm)",y = "Weight (grams)") 
  
# 1.2) Weight ~ ln(Final_Length): level-log model without scaling

ggplot( fish_data , aes(x = Final_Length , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Ln(Length) (cm)", y = "Weight (grams)") +
  scale_x_continuous(trans = log_trans(), breaks = c(10,20,30,40,50,60, 70))

# 1.3) ln(Weight) ~ Final_Length: log-level model without scaling

ggplot( fish_data , aes(x = Final_Length , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Length (cm)", y = "ln(Weight) (grams)") +
  scale_y_continuous( trans = log_trans())

# 1.4) ln(Weight) ~ ln(Final_Length): log-level model without scaling

ggplot( fish_data , aes(x = Final_Length , y = Weight)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "ln(Length) (cm)", y = "ln(Weight) (grams)") +
  scale_y_continuous( trans = log_trans()) +
  scale_x_continuous(trans = log_trans(), breaks = c(10,20,30,40,50,60, 70))

# Creating the new variables - Log Transformation

fish_data <- fish_data %>% mutate(ln_Weight = log(Weight),
                    ln_Final_Length = log(Final_Length))

### Check homoskedastic - Breusch-Pagan Test
lvl_lvl <- lm( Weight ~ Final_Length , data = fish_data)
log_lvl <- lm( ln_Weight ~ Final_Length , data = fish_data )
lvl_log <- lm( Weight ~ ln_Final_Length , data = fish_data )
log_log <- lm( ln_Weight  ~ ln_Final_Length , data = fish_data )

bptest( lvl_lvl ) # small p-value can´t reject the null hypothesis (homoscedasticity)
bptest( lvl_log ) # small p-value can´t reject the null hypothesis (homoscedasticity) 
bptest( log_lvl ) # big p-value can reject the null hypothesis (homoscedasticity) 
bptest( log_log ) # big p-value can reject the null hypothesis (homoscedasticity)

log_lvl <- lm_robust( ln_Weight ~ Final_Length , data = fish_data , se_type = "HC2" )
log_log <- lm_robust( ln_Weight  ~ ln_Final_Length  , data = fish_data , se_type = "HC2" )

# Model Comparison - Log Transformation Options

data_out <- "D:/CEU/Data_Analysis_2/DA2_Final_Project/out/"
htmlreg(list(lvl_lvl , log_lvl , lvl_log,  log_log ),
         type = 'html',
         custom.model.names = c("lvl_lvl", "log_lvl","lvl_log",
                                "log_log"),
         caption = "Log Transformations Comparison",
         file = paste0( data_out ,'logtrans_Weight_Length.html'), include.ci = FALSE)


rm(log_log,log_lvl,lvl_log,lvl_lvl)

```

### Weight and Width

```{r, echo=FALSE}

# 3) Weight and Width

# Now that it´s known how to treat the Length and Height variables, it's time to focus on the Width

# 3.1) ln(Weight) ~ Width: log-level model without scaling

ggplot( fish_data , aes(x = Height , y = Width)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "Width (%)", y = "ln(Weight) (grams)") +
  scale_y_continuous(trans = log_trans())

# 3.2) ln(Weight) ~ ln(Width): log-level model without scaling

ggplot( fish_data , aes(x = Height , y = Width)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = "ln(Width) (%)", y = "ln(Weight) (grams)") +
  scale_y_continuous(trans = log_trans()) +
  scale_x_continuous(trans = log_trans())

# Creating new variables - Log transformation

fish_data <- fish_data %>% mutate(ln_Width = log(Width))

### Check homoskedastic - Breusch-Pagan Test
log_lvl <- lm( ln_Weight  ~ Width  , data = fish_data )
log_log <- lm( ln_Weight  ~ ln_Width , data = fish_data )

bptest( log_lvl ) # small p-value can´t reject the null hypothesis (homoscedasticity)
bptest( log_log ) # small p-value can´t reject the null hypothesis (homoscedasticity)

log_lvl <- lm_robust( ln_Weight ~ Width , data = fish_data , se_type = "HC2" )
log_log <- lm_robust( ln_Weight  ~ ln_Width  , data = fish_data , se_type = "HC2" )

data_out <- "D:/CEU/Data_Analysis_2/DA2_Final_Project/out/"
htmlreg(list( log_lvl ,  log_log ),
        type = 'html',
        custom.model.names = c("log_lvl", "log_log"),
        caption = "Log Transformations Comparison",
        file = paste0( data_out ,'logtrans_Weight_Width.html'), include.ci = FALSE)

# Conclusion: 

# The analysis of the graphs and the comparisons between the models with and without the log-transformations,
# didn't show any significant statistic difference. The pattern expressed in both graphs are quite similar,
# And the results obtained in terms of R-squared are pretty close as well. 
# With that, due to the ease and more tangible interpretation of the results, the model chosen was the log-lvl one.

# Another important aspect to highlight is the fact that 
# the pattern observed can possibly fit a quadratic or cubic polynomial models.

# Evaluate the polynomial transformations for the variable Width

ggplot( data = fish_data, aes( x = Width, y = Weight ) ) + 
  geom_point( color='blue') +
  geom_smooth( formula = y ~ poly(x,2) , method = lm , color = 'red' )

ggplot( data = fish_data, aes( x = Width, y = Weight ) ) + 
  geom_point( color='blue') +
  geom_smooth( formula = y ~ poly(x,3) , method = lm , color = 'red' )

# Creating the new variables - Cubic transformation

fish_data <- fish_data %>% mutate(Width_sq = Width^2,
                                  Width_cb = Width^3)

# Comparing the regressions model with and without the quadratic/cubic transformations

reg5 <- lm_robust( ln_Weight  ~ ln_Final_Length + Height + Width , data = fish_data , se_type = "HC2" )
reg6 <- lm_robust( ln_Weight ~ ln_Final_Length + Height + Width + Width_sq , data = fish_data)
reg7 <- lm_robust( ln_Weight ~ ln_Final_Length + Height + Width + Width_sq + Width_cb , data = fish_data)

# Model Comparison

data_out <- "D:/CEU/Data_Analysis_2/DA2_Final_Project/out/"
htmlreg( list(reg1 , reg2 , reg5, reg6 , reg7 ),
         type = 'html',
         custom.model.names = c("Only Length", "Length + Height",
                                "Length + Height + Width", "Length + Height + Quadratic Width",
                                "Length + Height + Cubic Width"),
         caption = "Modelling the size of fishes by Length, Height and Width",
         file = paste0( data_out ,'model_comparison_Length_Height_Width.html'), include.ci = FALSE)

# Conclusion 

# Although the cubic and quadratic models showed a slightly better fit than the log-lvl,
# the interpretation is much easier under the simple model. 
# The differences aren't that significant enough in terms R-Squared as well.

# Remove non-used variables

rm(log_lvl,log_log)

```















